---
title: "GETurbineCaseStudyPart1"
output: html_document
---

__Read the data first__

```{r}

filenameToRead<-"train.csv"
turbineTrainData<-read.csv(filenameToRead)
str(turbineTrainData)

```


__Performing some data explorations__

```{r}
#Find the summary of the values

summary(turbineTrainData$X189)

summary(turbineTrainData$X190)

```
The summary clearly shows that the X190 and X189 have similar distribution

Now drawing the box plts to see the disribution

```{r}
#boxplots

boxplot(turbineTrainData$X189,turbineTrainData$X190,names=c("turbine 189","turbine 190"),main="Box Plots comparing turbine 189 & 190")

abline(a=median(turbineTrainData$X189),b=0,col="red",lty=2)

#boxplot is showing that the distribution of the datapoints in the quantiles is similar. The median is almost the same which indicates that X189 can be a good distribution for making predictions for X190

#inter quartile ranges
IQR(turbineTrainData$X189, na.rm = FALSE, type = 7)
IQR(turbineTrainData$X190, na.rm = FALSE, type = 7)
```

Now drawing the histogram
```{r}
#hist
par(mfrow=c(2,2))
hist(turbineTrainData$X189,main="Histogram for turbine189",xlab ='Power Values for 189')
density189<-density(turbineTrainData$X189)
plot(density189, main="Density for turbine189",xlab ='Power Values for 189')
polygon(density189, col="red", border="blue")

hist(turbineTrainData$X190,main="Histogram for turbine190",xlab ='Power Values for 190')
density190<-density(turbineTrainData$X190)
plot(density190, main="Density for turbine190",xlab ='Power Values for 190')
polygon(density190, col="red", border="blue")

```

the distribution is kind of skewed towards left 

correlations
```{r}

par(mfrow=c(1,1))
cor(turbineTrainData$X189,turbineTrainData$X190)

#plot the scatterplot
plot(turbineTrainData$X189,turbineTrainData$X190,xlab="Power values turbine 189", ylab="Power values turbine 190", main="Scatter Plot between turbine 190 and turbine 189")

```

So the above graphs indicate that there is strong relation between the two set of data points and it is possible to fit a line that passes through the bulk of points  
Also the histograms show that the distribution of 189 and 190 is similar and not Normal distribution


__Now fitting the model__
```{r}
linreg1=lm(X190~X189,data=turbineTrainData)
summary(linreg1)

#plot the points with the linear regression line
plot(turbineTrainData$X189,turbineTrainData$X190,xlab="Power values turbine 189", ylab="Power values turbine 190", main="Scatter Plot between turbine 190 & 189 with regression line",col="grey")
abline(lm(turbineTrainData$X190~turbineTrainData$X189), col="red")


#getting the residuals values using the summary function
residualsBySummary=linreg1$residuals

#calculate the SSE
trainSSE=sum(linreg1$residuals^2)
trainRMSE=sqrt(trainSSE/nrow(turbineTrainData))
trainRMSE
trainSST=sum((turbineTrainData$X190-mean(turbineTrainData$X190))^2)
trainR2=1-trainSSE/trainSST
trainR2

#another way to calculate
trainR2<-summary(linreg1)$adj.r.squared
trainMSE<-mean((turbineTrainData$X190 -linreg1$fitted.values)^2)
trainRMSE<-sqrt(trainMSE)
```

__Plot to overlap observed and fitted value__
```{r}
par(xpd=T, mar=par()$mar+c(0,0,0,6))
plot(turbineTrainData$X190,col="grey",main="Overlap of observed and fitted values on train data")
points(linreg1$fitted.values,col="red")

abline(h=1450,col="blue" )

numHighValues<-nrow(subset(turbineTrainData,turbineTrainData$X190>1450))
percentHighValues<-numHighValues/nrow(turbineTrainData)*100
percentHighValues

legend(1100,1500,c("Actual", "Fitted"), pch = c(1,1), col = c('grey','red'))

# Restore default clipping rect
par(mar=c(5, 4, 4, 2) + 0.1)

```

We see that about 7.5% points are not well estimated by the model


__Performing some linear regression diagnostics on the above model__
```{r}
#diagnostics
plot(linreg1, col="grey")
```

We can see that the diagnostic graphs show that the model does not have the issues:  
* non-linearity
* error between the actual and fitted value is normally distributed
* no Heteroscedasticity

Run the model on the test data to see the predictions
```{r}
#read the test data
turbineTestData<-read.csv("test.csv")
summary(turbineTestData)

predictions=predict(linreg1,newdata=turbineTestData)

head(predictions)
str(turbineTestData)

testErrorDiff=turbineTestData$X190-predictions
testSSE=sum(testErrorDiff^2)

testRMSE=sqrt(testSSE/nrow(turbineTestData))
testRMSE


#alternatively
testMSE<-mean((turbineTestData$X190 -predictions)^2)
testRMSE<-sqrt(testMSE)

#mean to be calculated on the trainig set
testErrorDiff=turbineTestData$X190-predictions
testSSE=sum(testErrorDiff^2)
testSST=sum((turbineTestData$X190-mean(turbineTestData$X190))^2)
testR2=1-testSSE/testSST
testR2

```
we can see that test R2 is slightly improved than the training R2. 


