---
title: "GETurbineCaseStudyPart2"
output: html_document
---

read the training data

```{r}
filenameToRead<-"train.csv"
turbineTrainData<-read.csv(filenameToRead)
str(turbineTrainData)

```

Performing some explorations

```{r}
#Find the summary of the values
summary(turbineTrainData)

```

Now drawing the box plts to see the distribution

```{r}
#boxplots
turbines<-names(turbineTrainData)
boxplot(turbineTrainData,names=turbines,main="Box plot comparing all turbines")
abline(a=median(turbineTrainData$X190),b=0,col="red",lty=2)
abline(a=summary(turbineTrainData$X190)[2],b=0,col="blue",lty=2)
abline(a=summary(turbineTrainData$X190)[5],b=0,col="blue",lty=2)


```
The box plot shows that for some of the turbines are producing the data similar to X190 while there are others which are not similar. So they might not have much impact while developing the model

Now checking the correlations

```{r}
cor(turbineTrainData)
```


__Generating models__

Regression Model1
using all the independent variables  

```{r}
model1=lm(X190~.,data=turbineTrainData)
summary(model1)

```


Regression Model2

After removing the not so important variables
```{r}
model2=lm(X190~X180+X186+X188+X189+X191+X193+X200,data=turbineTrainData)
summary(model2)

```

