---
title: "GETurbine"
output: html_document
---
This is the solution for the first question
•  It is recommended to first analyze the performance of turbine 190 compared to neighboring turbine 189.  Consider the error if turbine 189 is used as an estimator of turbine 190.  How well does this estimator perform?


__Read the data first__

```{r}

filenameToRead<-"train.csv"
turbineTrainData<-read.csv(filenameToRead)
str(turbineTrainData)

```

__Performing some data explorations__

```{r}
#Find the summary of the values

summary(turbineTrainData$X189)

summary(turbineTrainData$X190)

```
The summary clearly shows that the X190 and X189 have similar distribution


Now drawing the box plts and the histograms to see the disribution

```{r}
#boxplots

boxplot(turbineTrainData$X189,turbineTrainData$X190,names=c("turbine 189","turbine 190"),main="Box Plots comparing turbine 189 & 190")

abline(a=median(turbineTrainData$X189),b=0,col="red",lty=2)

#boxplot is showing that the distribution of the datapoints in the quantiles is similar. The median is almost the same which indicates that X189 can be a good distribution for making predictions for X190

#inter quartile ranges
IQR(turbineTrainData$X189, na.rm = FALSE, type = 7)
IQR(turbineTrainData$X190, na.rm = FALSE, type = 7)

#hist
par(mfrow=c(2,2))
hist(turbineTrainData$X189,main="Histogram for turbine189",xlab ='Power Values for 189')
density189<-density(turbineTrainData$X189)
plot(density189, main="Density for turbine189",xlab ='Power Values for 189')
polygon(density189, col="red", border="blue")

hist(turbineTrainData$X190,main="Histogram for turbine190",xlab ='Power Values for 190')
density190<-density(turbineTrainData$X190)
plot(density190, main="Density for turbine190",xlab ='Power Values for 190')
polygon(density190, col="red", border="blue")


#the distribution is kind of skewed towards left 

#correlations
cor(turbineTrainData$X189,turbineTrainData$X190)

#plot the points with the linear regression line
plot(turbineTrainData$X189,turbineTrainData$X190,xlab="Power values turbine 189", ylab="Power values turbine 190", main="Scatter Plot between turbine 190 and turbine 189")
#abline(lm(turbineTrainData$X190~turbineTrainData$X189), col="red")


```

So the above graphs indicate that there is strong relation between the two set of data points and it is possible to fit a line that passes through the bulk of points  
Also the histograms show that the distribution of 189 and 190 is similar and not Normal distribution


__Now fitting the model__
```{r}
linreg1=lm(X190~X189,data=turbineTrainData)
summary(linreg1)

#plot the points with the linear regression line
plot(turbineTrainData$X189,turbineTrainData$X190,xlab="Power values turbine 189", ylab="Power values turbine 190", main="Scatter Plot between turbine 190 & 189 with regression line",col="grey")
abline(lm(turbineTrainData$X190~turbineTrainData$X189), col="red")


fittingTrain=predict(linreg1)

#getting the residuals values using the summary function
residualsBySummary=linreg1$residuals

#calculating the residuals using the predictions and the actual value in teh training data
residualTrain=turbineTrainData$X190-fittingTrain

#both should be same
diff=residualTrain-residualsBySummary

#calculate the SSE
trainSSE=sum(linreg1$residuals^2)
trainRMSE=sqrt(trainSSE/nrow(turbineTrainData))
trainRMSE
trainSST=sum((turbineTrainData$X190-mean(turbineTrainData$X190))^2)
trainR2=1-trainSSE/trainSST
trainR2

#another way to calculate
trainMSE<-mean((turbineTrainData$X190 -linreg1$fitted.values)^2)
trainRMSE<-sqrt(MSE)


```

__Draw the standard error lines__ 
```{r}
par(mfrow=c(1,1))
preds<-predict(linreg1,se=TRUE)
#get the standard errors at any xo for the curve
se.bands=cbind(preds$fit+2*preds$se.fit,preds$fit-2*preds$se.fit)

#now plot the main region
plot(turbineTrainData$X189,turbineTrainData$X190,col="darkgrey")
#this is the main line
lines(turbineTrainData$X189,preds$fit,lwd=2,col="blue")
#now draw the SE lines
matlines(turbineTrainData$X189,se.bands,col="red",lty=2)


```


__Performing some linear regression diagnostics on the above model__
```{r}

par(mfrow=c(2,2))
#diagnostics
plot(linreg1,col="grey")


par(mfrow=c(1,1))
#plotting the outliers
plot(fitted.values(linreg1),rstudent(linreg1))
abline(h=c(-3,3), col="red")
which(rstudent(linreg1)>3)

#another way to get the outliers
library(car)
outliers=outlierTest(linreg1)
outliers

#get the high lebvergare points
highLevPoints<-influencePlot(linreg1)
highLevPoints
```
  
__Some obeservations__  
* We dont see many outliers or high leverage points
* The plot between the residuals and fitted values is almost linear which indicates that the linear regression model is a good fit

Suppose we try to use a polynomial degree of the X189 just to see if that improves the R2 significantly
```{r}

polyLin<-lm(X190~poly(X189,3),data=turbineTrainData)
summary(polyLin)

```
The summary shows that the power of 1 is only significant and other powers are not that significant. So we dont require a model which has nonlinear relationship with predictor variables.



Run the model on the test data to see the predictions
```{r}
#read the test data
turbineTestData<-read.csv("test.csv")
summary(turbineTestData)

turbineSubTestData=turbineTestData[c(1:2),]
predictionData=predict(linreg1,newdata=turbineTestData,interval="prediction")
predictions<-predictionData[1:nrow(predictionData)]
#coeffs<-coef(linreg1)
#predictions<-coeffs[1]+coeffs[2]*turbineTestData$X189
head(predictions)
str(turbineTestData)
testErrorDiff=turbineTestData$X190-predictions

testSSE=sum(testErrorDiff^2)
testRMSE=sqrt(testSSE/nrow(turbineTestData))
testRMSE
#mean to be calculated on the trainig set
testSST=sum((turbineTestData$X190-mean(turbineTestData$X190))^2)
testR2=1-testSSE/testSST
testR2
```
we can see that test R2 is slightly improved than the training R2. 

__see how the predcitions and actual data is plotted__
```{r}
plot(turbineTestData$X190)
points(predictions,col="red")
```



trying with the log transformation
```{r}
xx<-lm(X190~X189+I(X189^2),data=turbineTrainData)
summary(xx)
plot(xx)



```
